### 2019-08-20 paper/paper.tex
paper/paper.tex
Reorder free param vector.
Most table drafts in and text rearranged.

\documentclass{article}

\usepackage{siunitx}
\usepackage{amsmath}

@@

\begin{center}
  \fontsize{14}{20}{\bf Expanded Optimization for Discovering Optimal Lateral
  Handling Bicycles}
\end{center}

\begin{center}
  \normalsize{\bf{Jason K. Moore, Mont Hubbard,  Ronald A. Hess}}
\end{center}

\begin{center}
  \begin{tabular}{c}
    Department of Mechanical and Aerospace Engineering\\
    University of California, Davis\\
    One Shields Avenue, Davis, CA, USA 95616\\
    e-mail: jkm@ucdavis.edu, mhubbard@ucdavis.edu, rahess@ucdavis.edu\\
  \end{tabular}
\end{center}

\appendix

\section{Free Parameters and Constraints}

\subsection{Rear Frame [D]}

There are 5 free parameters associated with the rear frame's inertial
characteristics:

\begin{description}
  \item[$m_D$] Mass of the rear frame.
  \item[$x_D$] Location along the $X$ axis of the rear frame center of mass.
  \item[$z_D$] Location along the $Z$ axis of the rear frame center of mass.
  \item[$k_{Dxx}$] Central radius of gyration with respect to the $X$ axis
    where $k_{Dxx}^2=I_{Dxx} / m_D$.
  \item[$k_{Dzz}$] Central radius of gyration with respect to the $Z$ axis
    where $k_{Dzz}^2=I_{Dzz} / m_D$.
  \item[$\alpha_D$] Angle between the $X$ axis and the primary principal
    inertia axis.
\end{description}

Assumptions:

\begin{itemize}
  \item The body is symmetric with respect to the $XZ$ plane, i.e. $I_{Dxy} =
    I_{Dyz} = 0$.
  \item We require that the center of mass not penetrate the ground, i.e. $z_D
    < 0$.
  \item We require that the mass be positive, i.e. $m_D > 0$.
  \item We allow for any angular orientation of the $XZ$ principal directions
    but limit the angle to $-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$.
  \item We enforce that the body is planar with $k_{Dyy} = \sqrt{k_{Dxx}^2 +
    k_{Dyy}^2}$.
\end{itemize}

@@

We assume that the person's joint configurations are such that they are in a
nominal configuration for pedaling, i.e. an average normal everyday riding
position. The person is assumed to be symmetric about the $XZ$ plane. There are
four fixed parameters that enforce this assumption:

@@

We allow the rider to be rotated about the $Y$ axis and positioned anywhere
within the plane of symmetry. Three free parameters are associated with these
assumptions:

\begin{description}
  \item[$\alpha_P$] Angle between the $X$ axis and the primary principal
    inertia axis of the person.
  \item[$x_P$] Location along the $X$ axis of the person's center of mass.
  \item[$z_P$] Location along the $Z$ axis of the person's center of mass.
\end{description}

To prevent the rider from being positioned and oriented such that their body is
not penetrating the the ground we introduce two dimensions that define a cross
whose apex is at the center of mass of the person and the cross axes are
parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance
along the principal axis to the tip of the toes and $w_P$ is the distance along
the second principal axes to the tip of the hands.

\begin{align}
  z_P + l_P / 2 \cos{\alpha_P} & \leq 0 \\
  z_P + w_P / 2 \sin{\alpha_P} & \leq 0 \\
  z_P - l_P / 2 \cos{\alpha_P} & \leq 0 \\
  z_P - w_P / 2 \sin{\alpha_P} & \leq 0
\end{align}

@@

\begin{description}
  \item[$m_H$] The mass of the rear frame.
  \item[$k_{Hxx}$] Central radius of gyration with respect to the $X$ axis
    where $k_{Hxx}^2=I_{Hxx} / m_H$.
  \item[$k_{Hyy}$] Central radius of gyration with respect to the $X$ axis
    where $k_{Hyy}^2=I_{Hyy} / m_H$.
  \item[$k_{Hzz}$] Central radius of gyration with respect to the $X$ axis
    where $k_{Hzz}^2=I_{Hzz} / m_H$.
  \item[$\alpha_H$] Angle between the $X$ axis and the primary principal
    inertia axis.
  \item[$x_B$] Location along the $X$ axis of the rear frame's center of mass.
  \item[$z_H$] Location along the $Z$ axis of the rear frame's center of mass.
\end{description}

Assumptions:

\begin{itemize}
  \item We allow for any angular orientation of the $XZ$ principal directions
    but limit the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$.
\end{itemize}

\subsection{Rear Wheel [R]}

We enforce the assumption that the wheel should be wheel-like, i.e. $I_{Ryy} =
m_R r_R^2$ and $I_{Rxx} = I_{Rzz} = I_{Ryy} / 2$ thus there are two free
parameters:

\begin{description}
  \item[$r_R$] Outer radius of the rear wheel.
  \item[$m_R$] Mass of the rear wheel.
\end{description}

We also enforce that the mass and radius should be positive with:

\begin{align}
  m_R \geq 0 \\
  r_R \geq 0
\end{align}

\subsection{Front Wheel}

Same as rear wheel.

\subsection{Total Bike}

There are three remaing free parameters associated with the fundamental
geometry of the bicycle:

\begin{description}
  \item[$w$] Wheelbase
  \item[$c$] Trail
  \item[$\lambda$] Steer axis tilt
\end{description}

To prevent the wheels from overlapping we introduce:

\begin{align}
  w > r_F + r_R
\end{align}

@@

We require that the bicycle be liftable by an average person with:

\begin{align}
  m_T = m_H + m_D + m_P + m_F + m_R \leq m_P + 25 \si{kg}
\end{align}

Finally, we require that the bicycle not topple forward during hard breaking
and backward during hard acceleration where:

\begin{align}
  -g < a < \frac{g}{4}
\end{align}

This translates to two constraints that bound the total center of mass in a triangle in teh $XZ$ plane:

\begin{align} \frac{x_T}{|z_T|} < \frac{1}{4} \\ \frac{x_T - w}{|z_T|} < -1
\end{align}

@@

\begin{align}
  \mathbf{p} =
  \begin{bmatrix}
    % front frame
    m_D \\
    x_D \\
    z_D \\
    k_{Dxx} \\
    k_{Dzz} \\
    \alpha_D \\
    % person
    x_P \\
    z_P \\
    \alpha_P \\
    % front frame
    m_H \\
    x_H \\
    z_H \\
    k_{Hxx} \\
    k_{Hyy} \\
    k_{Hzz} \\
    I_{Hxz} \\
    \alpha_H \\
    % rear wheel
    r_R \\
    m_R \\
    % front wheel
    r_F \\
    m_F \\
    % primary geometry
    w \\
    c \\
    \lambda
  \end{bmatrix}
\end{align}

@@

>>>>

\documentclass{bmd2019a}

\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{booktabs}

@@

\begin{flushleft}
{\fontsize{16pt}{20pt}\selectfont%
  Expanded Optimization for Discovering\\}
{\fontsize{16pt}{20pt}\selectfont%
  Optimal Lateral Handling Bicycles}
\end{flushleft}

%%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
\begin{flushleft}
  {\fontsize{12}{14}{Jason K. Moore, Mont Hubbard, and Ronald A. Hess}\\}
  \textit{Mechanical and Aerospace Engineering\\
          University of California, Davis\\
          One Shields Avenue, Davis, CA, USA 95817\\
          e-mail: jkm@ucdavis.edu, mhubbard@ucdavis.edu, rahess@ucdavis.edu}
\end{flushleft}

\section*{Abstract}
%
Physical design features of ground vehicles can affect their lateral handling
qualities. Geometry, mass, and mass distribution of the vehicle's primary
components as well as tire characteristics are primary contributors to poor and
good handling due to their important influence on the vehicle's dynamics. In
past work, we have presented a theoretical and computational framework for
assessing the lateral task-independent handling qualities of simplified single
track vehicle designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed
that minimizing our handling quality metric (HQM) can produce theoretically
optimal handling designs when only four geometric parameters are explored as
the optimization variables~\cite{Moore2016}. The present work's goal is to
expand this optimization problem to 24 geometry, mass, and inertial parameters
as the optimization variables so that a broader search space of optimal bicycle
designs is considered. This task is complicated by the fact that we desire to
limit the search space to physically realizable designs. Doing so introduces 35
linear and non-linear constraint equations on the optimization variables. We
formulate a constrained optimization problem and use nonlinear programming to
discover optimal, yet realizable, bicycle designs.

\section*{Introduction}
%
We introduced a method of optimizing four geometric parameters of a bicycle's
design to minimize the so called Handling Quality Metric
(HQM)~\cite{Moore2016}. Here we expand that method to optimize all X geometric
and inertial parameters of the benchmark parameterimization of the linear
Whipple-Carvallo bicycle model under constraints that guarantee a physically
realizable bicycle. This improves over the prior work by expanding the search
space with many more parameters and the guarntee of realizability.

\section*{Problem Definition}

We search for 24 optimization variables that minimize the Handling Quality
Metric defined in ~\cite{Hess2012}.

Our formulation exposes 24 optimization variables
$\mathbf{p}\in\mathbb{R}^{24}$ that are both reformulations of and additions to
the parameters of the benchmark bicycle~\cite{Meijaard2007}. We separate the
rider's inertial parameters from the rear frame of the bicycle for the purposes
of rider/bicycle relative configuration but do not introduce any more than the
benchmark's four rigid bodies. The variables are as follows;
%
\begin{itemize}
  \itemsep-0.25em
  \item wheelbase, trail, steer axis tilt, wheel radii
  \item somersault angle of the person (same direction as bicycle pitch)
  \item masses of the front frame, rear frame, front wheel, and rear wheel
  \item mass center locations of the front frame, rear frame, and rider
  \item principal central radii of gyration of front and rear frames
  \item principal inertia axis pitch angles of the front frame, rear frame, and rider
\end{itemize}

We formulate bounds on the variables as well as equality and inequality
constraints among the variables to ensure a physically realizable bicycle.
Below the constraints are defined and grouped by the associated rigid body or
collection thereof:
%
\begin{description}

  \item[Vehicle] The resulting combination of the five rigid bodies.
    \begin{itemize}
      \itemsep0em
      \item The physical extents of the bicycle and rider must exist above the
        ground plane.
      \item Both bicycle and rider are symmetric about the sagittal
        plane.
      \item The mass of each bicycle rigid body is positive, greater than a
        minimum value, and the total mass is below a reasonably lift-able
        amount.
      \item The wheels cannot overlap, i.e. wheelbase > front radius + rear
        radius.
      \item The bicycle cannot topple forward during hard braking or backward
        during hard acceleration, i.e. -1 g < acceleration < 1 g.
    \end{itemize}

  \item[Wheels] Both front and rear wheels have identical constraints.
    \begin{itemize}
      \itemsep0em
      \item Wheel radius must be greater than a minimum value.
      \item Wheels are inertially wheel-like, i.e.  $I_{rot}=mr^2,
        I_{rad}=I_{rot}/2$.
    \end{itemize}

  \item[Frames] Both front and rear frames have identical constraints.
    \begin{itemize}
      \itemsep0em
      \item A pair of principal directions lie in the roll/yaw
        plane to maintain symmetry and one of these axes is within $\pm \pi/4$
        rad of the ground plane.
      \item The mass and inertia of the frames are positive and large enough to
        be constructed from materials in a space frame of specified minimal
        density.
      \item The frames are planar in nature, i.e.
        $k_{lateral}^2=k_{longitudinal}^2+k_{vertical}^2$, where $k$ is the
        radius of gyration.
    \end{itemize}

  \item[Rider] A single rigid body represents the rider.
    \begin{itemize}
      \itemsep0em
      \item Rider mass is that of an average person.
      \item The rider's joint angles are fixed in a nominal configuration
        typical of upright bicycling and the resulting mass distribution is
        derived from standard body segment estimation methods.
      \item One rider principal axis is parallel with the bicycle pitch axis to
        enforce symmetry.
    \end{itemize}

\end{description}

The constraint statements above translate into upper and lower bounds on 12 of
the 24 variables $\mathbf{p}^U,\mathbf{p}^L$ and 14 linear and 21 nonlinear
constraint $\mathbf{g}(\mathbf{p})$ among the variables and thus can be
expressed in the standard non-linear programming form.
%
\begin{equation}
  \begin{aligned}
    & \underset{\mathbf{p}}{\text{minimize}} & & max(\textrm{HQM}(\mathbf{p})) \\
    & \text{subject to} & & \\
    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
    & & & \mathbf{g}^L \leq \mathbf{g}(\mathbf{p}) \leq \mathbf{g}^U \\
  \end{aligned}
\end{equation}

\subsection*{Parameters}
%
We introduce a new parameterization of the linear Whipple-Carvallo bicycle
model. This parameterization differs from the benchmark
parameterization~\cite{Meijaard2007} in that the person and rear frame are
separate rigid bodies each with their on inertial parameters and that the
inertial parameters of all bodies are expressed in terms of principal central
radii of gyration so that the mass is decoupled from the inertia. We also
introduce two simple dimensional parameters that define the geometric extents
of the person. Table~\ref{tab:parameters} provides the parameters and the
initial values which are derived from.

% TODO : Create a table of the parameters and values.
\begin{table}
  \caption{Parameters and their default values.}
  \begin{tabular}{llll}
    \toprule
    Variable & Value & Units & Description \\
    \midrule
    $c$ &  0.068581 & \si{\meter} & Trail \\
    $w$ &  1.1210 & \si{\meter} & Wheelbase \\
    $\lambda$ &  0.39968 & \si{\radian} & Steer axis tilt \\
    $g$ &  9.81 & \si{\meter\per\second\squared} & Acceleration due to gravity \\
    $v$ &  3.0 & \si{\meter\per\second} & Forward speed \\
    Rear Wheel [R] & & \\
    \midrule
    $m_R$ &  3.11 & \si{\kilogram} & mass \\
    $r_R$ &  0.34096 & \si{\meter} & Rear wheel radius \\
    $x_R$ & 0 & \si{\meter} & Rear wheel mass center \\
    $y_R$ & 0 & \si{\meter} & Rear wheel mass center \\
    $z_R$ & -0.34096 & \si{\meter} & Rear wheel mass center \\
    $k_{Raa}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration \\
    $k_{Rbb}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration  \\
    $k_{Ryy}$ &  0.22136 & \si{\meter} & Rear wheel central principal radii of gyration \\
    Front Wheel [F] & & \\
    \midrule
    $m_F$ &  2.02 & \si{\kilogram} & Mass \\
    $r_F$ &  0.34353 & \si{\meter} & Radius \\
    $x_F$ &  1.1210 & \si{\meter} & Mass center \\
    $y_F$ & 0.0 & \si{\meter} & Mass center \\
    $z_F$ & -0.34353 & \si{\meter} & Mass center \\
    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radius of gyration \\
    \midrule
    $l_P$ &  1.7280 & \si{\meter} & Body length \\
    $w_P$ &  0.48300 & \si{\meter} & Body width \\
    $m_P$ &  83.500 & \si{\kilogram} & Mass \\
    $x_P$ &  0.31577 & \si{\meter} & Mass center \\
    $y_P$ & 0.0 & \si{\meter} & Mass center \\
    $z_P$ & -1.0990 & \si{\meter} & Mass center \\
    $k_{Paa}$ &  0.36797 & \si{\meter} & Central principal radius of gyration \\
    $k_{Pbb}$ &  0.15276 & \si{\meter} & Central principal radius of gyration \\
    $k_{Pyy}$ &  0.36717 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_P$ & 0.18618 & \si{\radian} & Principal axis angle \\
    \midrule
    $m_H$ & 3.2200 & \si{\kilogram} & Mass \\
    $x_H$ & 0.86695 & \si{\meter} & Mass center \\
    $y_H$ & 0.0 & \si{\meter} & Mass center \\
    $z_H$ & -0.74824 & \si{\meter} & Mass center \\
    $k_{Haa}$ & 0.29556 & \si{\meter} & Central principal radius of gyration \\
    $k_{Hbb}$ & 0.14493 & \si{\meter} & Central principal radius of gyration \\
    $k_{Hyy}$ & 0.27630 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_H$ & 0.36995 & \si{\radian} & Principal axis angle \\
    \midrule
    $m_D$ &  9.8600 & \si{\kilogram} & Mass \\
    $x_D$ &  0.27595 & \si{\meter} & Mass center \\
    $y_D$ & 0 & \si{\meter} & Mass center \\
    $z_D$ & -0.53784 & \si{\meter} & Mass center \\
    $k_{Daa}$ &  0.28587 & \si{\meter} & Central principal radius of gyration \\
    $k_{Dbb}$ &  0.22079 & \si{\meter} & Central principal radius of gyration \\
    $k_{Dyy}$ &  0.36539 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_D$ &  1.1722 & \si{\radian} & Principal axis angle \\
    \bottomrule
  \end{tabular}
\end{table}

These parameters can be transformed into the benchmark parameterization
r adily, but not vice versa.

We search for the optimal values of a subset of 23 of the principal parameters.

\begin{align}
  \mathbf{p} =
    [w \quad
     c \quad
     \lambda \quad
     m_D \quad
     x_D \quad
     z_D \quad
     k_{Daa} \quad
     k_{Dbb} \quad
     \alpha_D \quad
     x_P \quad
     z_P \quad
     \alpha_P \quad
     m_H \quad
     x_H \quad
     z_H \quad
     k_{Haa} \quad
     k_{Hbb} \quad
     k_{Hyy} \quad
     \alpha_H \quad
     r_R \quad
     m_R \quad
     r_F \quad
     m_F]
\end{align}

\subsection*{Constraints}
%
The optimal principal parameters are subject to a set of constraints designed
to ensure a physically realizable bicycle is obtained.

\begin{table}
  \caption{Inequality constraints}
  \begin{tabular}{lll}
    \toprule
    Constraint & Equation & Description \\
    \midrule
    $c_1$    & $\sqrt{I_{Hxx}^2+I_{Hzz}^2} \geq I_{Hyy}$           & Narrow front frame. \\
    $c_2$    & $0 \geq z_P + l_P/2 \cos{\alpha_P}$                 & Person cannot penetrate ground. \\
    $c_3$    & $0 \geq z_P + w_P/2 \sin{\alpha_P}$                 & Person cannot penetrate ground. \\
    $c_4$    & $0 \geq z_P - l_P/2 \cos{\alpha_P}$                 & Person cannot penetrate ground. \\
    $c_5$    & $0 \geq z_P - w_P/2 \sin{\alpha_P}$                 & Person cannot penetrate ground. \\
    $c_6$    & $x_T \geq |z_T|/4$                                  & Maximum acceleration of $1/4g$. \\
    $c_7$    & $w-x_T \geq 3/4|z_T|$                               & Maximum deceleration of $3/4g$. \\
    $c_8$    & $2k_{Hyy} \geq \sqrt{(x_H - w)^2 + (z_H + r_F)^2}$  & Minimal inertial spread. \\
    $c_9$    & $2k_{Dyy} \geq \sqrt{(x_D - 0)^2 + (z_D + r_R)^2}$  & Minimal inertial spread. \\
    $c_{10}$ & $w \geq r_F +r_R$                                   & Non-overlapping wheels. \\
    $c_{11}$ & $25\si{\kg} \geq m_D + m_H + m_R + m_F$             & Maximum bicycle mass. \\
    $c_{12}$ & $-z_D \geq 1.4 k_{Dyy}$                             & Rear frame cannot penetrate ground. \\
    $c_{13}$ & $-z_H \geq 1.4 k_{Hyy}$                             & Front frame cannot penetrate ground. \\
    $c_{14,\ldots,21}$ & $0 \geq s_1,\ldots,s_8$                   & Closed loop stability. \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}
  \caption{Equality constraints}
  \begin{tabular}{lll}
    \toprule
    Constraint & Equation & Description \\
    \midrule
    $g_1$ & $I_{Dyy} = \sqrt{I_{Dxx}^2 + I_{Dzz}^2}$ & Rear frame is planar. \\
    $g_2$ & $k_{Ryy} = r_R$  & Rear wheel is a ring \\
    $g_3$ & $k_{Raa} = k_{Ryy}/2$ & Rear wheel is a ring \\
    $g_4$ & $k_{Rbb} = k_{Ryy}/2$ & Rear wheel is a ring \\
    $g_5$ & $k_{Fyy} = r_F$  & Front wheel is a ring \\
    $g_6$ & $k_{Faa} = k_{Fyy}/2$ & Front wheel is a ring \\
    $g_7$ & $k_{Fbb} = k_{Fyy}/2$ & Front wheel is a ring \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}
  \caption{Parameter Bounds}
  \begin{tabular}{ccccc}
    \toprule
    Min & & Parmeter & & Max\\
    \midrule
    $-\infty$ & $\leq$ & $w       $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $c       $  &  $\leq$  &  $\infty$ \\
    $-\pi/2 $ & $\leq$ & $\lambda $  &  $\leq$  &  $\pi/2$ \\
    $1.0    $ & $\leq$ & $m_D     $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $x_D     $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $z_D     $  &  $\leq$  &  $0.0$    \\
    $0.0    $ & $\leq$ & $k_{Daa} $  &  $\leq$  &  $\infty$ \\
    $0.0    $ & $\leq$ & $k_{Dbb} $  &  $\leq$  &  $\infty$ \\
    $-\pi/2 $ & $\leq$ & $\alpha_D$  &  $\leq$  &  $\pi/2$  \\
    $-\infty$ & $\leq$ & $x_P     $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $z_P     $  &  $\leq$  &  $0.0$    \\
    $-\pi/2 $ & $\leq$ & $\alpha_P$  &  $\leq$  &  $\pi/2$ \\
    $0.25   $ & $\leq$ & $m_H     $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $x_H     $  &  $\leq$  &  $\infty$ \\
    $-\infty$ & $\leq$ & $z_H     $  &  $\leq$  &  $0.0$ \\
    $0.0    $ & $\leq$ & $k_{Haa} $  &  $\leq$  &  $\infty$ \\
    $0.0    $ & $\leq$ & $k_{Hbb} $  &  $\leq$  &  $\infty$ \\
    $0.0    $ & $\leq$ & $k_{Hyy} $  &  $\leq$  &  $\infty$ \\
    $-\pi/2 $ & $\leq$ & $\alpha_H$  &  $\leq$  &  $\pi/2$ \\
    $0.127  $ & $\leq$ & $r_R     $  &  $\leq$  &  $\infty$ \\
    $1.0    $ & $\leq$ & $m_R     $  &  $\leq$  &  $\infty$ \\
    $0.127  $ & $\leq$ & $r_F     $  &  $\leq$  &  $\infty$ \\
    $1.0    $ & $\leq$ & $m_F     $  &  $\leq$  &  $\infty$ \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection*{Rear Frame [D]}

There are five free parameters associated with the rear frame's inertial
characteristics:

\begin{description}
  \item[$m_D$] Mass of the rear frame.
  \item[$x_D$] Location along the $X$ axis of the rear frame center of mass.
  \item[$z_D$] Location along the $Z$ axis of the rear frame center of mass.
  \item[$k_{Daa}$] Maximum principal central radius of gyration in the $XZ$
    plane.
  \item[$k_{Dbb}$] Minimum principal central radius of gyration in the $XZ$
    plane.
  \item[$\alpha_D$] Angle between the $X$ axis and the maximum principal axis
    that lies in the $XZ$ plane.
\end{description}

Several constraints are set for the rear frame. We constrain the rear frame to
be planar, $g_1$ and symmetric with respect to the $XZ$ plane, i.e. $I_{Dxy} =
I_{Dyz} = 0$. We prevent the rear frame from penetrating the ground by limiting
the inertial spread with respect to its mass center, $c_{12}$, but also set a
minimum inertial spread to ensure a frame can span from the rear wheel to the
mass center of the rear frame, $c_9$. The spread factor in $c_{12}$ of 1.4 is
based on the ratio of geometrical spread of a typical bicycle frame and its
radius of gyration. Several parameters are bounded. We require the rear frame
mass to be positive, the center of mass not penetrate the ground, and we allow
for any angular orientation of the $XZ$ principal directions but limit the
angle to $-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$.

@@

We assume that the person's joint configurations are such that they are in a
nominal configuration for pedaling, i.e. an average normal everyday riding
position on a typical bicycle. The person is assumed to be symmetric about the
$XZ$ plane. We allow the rider to be rotated about the $Y$ axis and positioned
anywhere within the plane of symmetry above the ground. Three free parameters
are associated with these assumptions:

\begin{description}
  \item[$x_P$] Location along the $X$ axis of the person's center of mass.
  \item[$z_P$] Location along the $Z$ axis of the person's center of mass.
  \item[$\alpha_P$] Angle between the $X$ axis and the primary principal
    inertia axis of the person.
\end{description}

There are four fixed parameters that enforce this assumption:

@@

To prevent the rider from being positioned and oriented such that their body is
not penetrating the the ground we introduce two dimensions that define a cross
whose apex is at the center of mass of the person and the cross axes are
parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance
along the principal axis to the tip of the toes and $w_P$ is the distance along
the second principal axes to the tip of the hands. Constraints $c_2,\ldots,c_5$.

@@

\begin{description}
  \item[$m_H$] The mass of the rear frame.
  \item[$x_B$] Location along the $X$ axis of the rear frame's center of mass.
  \item[$z_H$] Location along the $Z$ axis of the rear frame's center of mass.
  \item[$k_{Haa}$] Central radius of gyration with respect to the $X$ axis
    where $k_{Hxx}^2=I_{Hxx} / m_H$.
  \item[$k_{Hyy}$] Central radius of gyration with respect to the $X$ axis
    where $k_{Hyy}^2=I_{Hyy} / m_H$.
  \item[$k_{Hbb}$] Central radius of gyration with respect to the $X$ axis
    where $k_{Hzz}^2=I_{Hzz} / m_H$.
  \item[$\alpha_H$] Angle between the $X$ axis and the primary principal
    inertia axis.
\end{description}

We allow for any angular orientation of the $XZ$ principal directions but limit
the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$.  We prevent the
rear frame from penetrating the ground by limiting the inertial spread with
respect to its mass center, $c_{13}$, but also set a minimum inertial spread to
ensure a frame can span from the rear wheel to the mass center of the rear
frame, $c_8$. The spread factor in $c_{13}$ of 1.4 is based on the ratio of
geometrical spread of a typical bicycle frame and its radius of gyration. The
front frame is not planar but is narrow with respect to the $XZ$ plane, which
is enforced by constraint $c_1$.

\subsection{Front [F] and Rear [R] Wheels}

We enforce the assumption that both wheels have moments of inertia of that of a
simple ring, $g_2 \ldots g_7$ and that mass and radius should be greater than
some minimal size based on small bicycle wheel.

\subsection{Total Bike [T]}

\subsubsection{Parameters}

There are three remaing parameters associated with the fundamental geometry of
the bicycle: wheelbase $w$, trail $c$, and steer axis tilt $\lambda$ all the
same as defined in~\cite{Meijaard2007}.

\subsubsection{Bounds}

@@

\subsubsection{Constraints}

We introduce a constraint $c_{10}$ that prevents the wheels from physically
overlapping and require that the bicycle be liftable by an average person,
$c_{11}$. Finally, we require that the bicycle not topple forward during hard
breaking or backward during hard acceleration.

\begin{align}
  -\frac{3g}{4} < \textrm{acceleration} < \frac{g}{4}
\end{align}

This translates to two constraints, $c_6,c_7$ that bound the total center of
mass $(x_T,z_T)$ in a triangle in the $XZ$ plane.

Lastly, we constrain the eight closed loop eigenvalues associated with the
controller in ~\cite{Hess2012} to be stable, i.e. have negative real parts.
These are expressed in constraints $c_{14},\ldots,c_{21}$. Closed loop
stability is required for the HQM to provide a meaningful result.

@@

\begin{align}
  \mathbf{p} =
  \begin{bmatrix}
    % primary geometry
    w \\
    c \\
    \lambda
    % front frame
    m_D \\
    x_D \\
    z_D \\
    k_{Dxx} \\
    k_{Dzz} \\
    \alpha_D \\
    % person
    x_P \\
    z_P \\
    \alpha_P \\
    % front frame
    m_H \\
    x_H \\
    z_H \\
    k_{Hxx} \\
    k_{Hyy} \\
    k_{Hzz} \\
    I_{Hxz} \\
    \alpha_H \\
    % rear wheel
    r_R \\
    m_R \\
    % front wheel
    r_F \\
    m_F \\
  \end{bmatrix}
\end{align}

@@

\bibliographystyle{acm}
\bibliography{BMD2019}

\section*{Solving}
\section*{Results}
\section*{Discussion and Conclusion}
